\revision{Confirmation of \srcfull{} as a long period radio transient}{A new long period radio transient: Discovery of pulses repeating every 1.16 hours from \srcfull{}}

### 2025-07-10 paper/main.tex
More tweaks to the paper courtesy of Natasha

\src{} was detected \revision{blindly}{} with the Murchison Widefield Array \citep[MWA;][]{Tingay2013} in observations taken between June--September 2024 under the `Galactic Plane Monitor', project code G0080. The prime objective of the campaign was to search for long-period transients by scanning the Galactic plane over $285^\circ < l < 65^\circ$, $|b| < 15^\circ$ on a bi-weekly cadence at 185--215\,MHz \citep[see Methods of][]{2023Natur.619..487H}; a full description will be released by Hurley-Walker et al., in prep. We also observed over June--September 2022 but made no detections in these data, as reported in \citetalias{2024MNRAS.535..909D}. \revision{}{The root-mean-squared (RMS) noise level in these observations on the processed four-second timescale is $\sim$80\,mJy\,beam$^{-1}$ in this region. A time-based convolution filter employed to extract pulse-like signals, as described by Horv\'ath et al., submitted, yielded the first detection in this data set of a pulse whose peak flux density was 665\,mJy.}
Subsequent to its redetection, we commenced a Director's Discretionary Time (DDT) campaign under project code D0042, to obtain further observations, adjusting the frequency from 200.32\,MHz to 184.96\,MHz for some of the observations to attempt to improve measurements of its DM.
Time of arrival (ToA) predictions for the DDT observations were made using a preliminary ephemeris derived from the GPM detections, which, by virtue at being at the same or similar frequencies, absorbed the delays due to dispersion and scattering into the `PEPOCH' term.

@@

We identified two faint pulses (4--5 $\sigma$; see Table~\ref{tbl:toas}) in the observation taken on 2024-10-04. However, observations taken on 2024-10-12 did not yield any pulses (at 3-$\sigma$ level). We inspected the data around the expected time of arrival (using Equation~\ref{eqn:timing}) but did not find any pulses. The instantaneous noise level, per 10\,s integration time bin, around the expected time of arrivals is close to 1\,mJy. If pulse-to-pulse flux density variations, similar to ones observed in MWA data (Table~\ref{tbl:toas}), also occur at L/S band, then even a factor 1.5--2 variation can be sufficient to explain the detection of pulses on 2024-10-04 and the absence of pulses on 2024-10-12, given that the source is faint at higher frequencies.

@@

For the MeerKAT pulse with two components, we fit only the brighter component.

@@

These ToAs were then barycentred and fit for both period and DM, resulting in the values given in \Tab~\ref{tbl:ephemeris}.
The residuals are presented in \Fig~\ref{fig:pulse_details} and \Tab~\ref{tbl:toas}, along with other fitted properties of the pulses.
\revision{We also tried fitting a timing model that included a spin-down parameter, and found that it was consistent with zero spin-down to within $1\sigma$, and so is not included in the ephemeris.}{We also tried fitting a timing model that included a spin-down/up parameter (fixing the DM to $733\,{\rm pc}\,{\rm cm}^{-3}$), and found that it was consistent with zero.
However, owing to recent discovery that CHIME/ILT\,J1634+44 is spinning up \citep[$\dot{P} = (-9.03 \pm 0.11) \times 10^{-12}\,{\rm s}\,{\rm s}^{-1}$;][]{2025arXiv250705139D,2025arXiv250705078B}, we report the fitted period derivative value of $(-10 \pm 92) \times 10^{-12}\,{\rm s}\,{\rm s}^{-1}$ instead of only providing an upper limit for the spin-down.}

@@

The variability of the pulse brightness, coupled with the fact that none of our detections were simultaneous at multiple telescopes, prohibit a direct spectral index measurement.
We note, however, that the fitted peak flux densities of the ASKAP and MeerKAT pulses are broadly consistent with the MWA flux densities at nearby epochs if an approximate spectral index of $\alpha = -2.0$ is assumed.
The bottom panel of \Fig~\ref{fig:pulse_details} shows the peak flux densities scaled to 1\,GHz under this assumption, revealing a relatively slow change of the pulses' brightness, not unlike GLEAM-X\,J1627$-$5235 \citep{2022Natur.601..526H}.
The in-band spectral index measurement of $\alpha=-3.1$ reported for the 2023 ASKAP pulse \citepalias{2024MNRAS.535..909D}, if consistent across the timespan of our observations, implies a turnover or spectral curvature at intermediate frequencies \citep[perhaps comparable with GPM\,1839$-$10][]{2023Natur.619..487H}.

@@

We conjecture that if \src{} is a WD pulsar, its intermittency may be a result of the spin-orbital resonance being near, but not exactly, a small-integer ratio, creating a aliased beating pattern on a time scale of months.
The same explanation may be applied to other LPTs with activity cycles that last on the order of months.
An alternative explanation is that the radio emission is drived by particle wind outflows from the main sequence companion.
Long-term monitoring of LPTs will reveal whether the intermittency is periodic, and thus help discriminate between these two scenarios.

@@

@@

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.48\linewidth]{J1755_optical.pdf}
    \includegraphics[width=0.48\linewidth]{J1755_nir.pdf}
    \caption{The $1\sigma$ positional uncertainty of \src{} derived from the MeerKAT observation taken on MJD 60602 is shown as the green ellipses overlaid on the optical (z, i, r) Panoramic Survey Telescope and Rapid Response System (PanSTARRS) image (left panel), and the infrared (K, H, J) United Kingdom Infrared Telescope Deep Sky Survey (UKIDSS) image (right panel).}
    \label{fig:localization}
\end{figure*}

>>>>

\revision{\src{} was detected blindly with the Murchison Widefield Array \citep[MWA;][]{Tingay2013} in observations taken between June--September 2024 under the `Galactic Plane Monitor', project code G0080. The prime objective of the campaign was to search for long-period transients by scanning the Galactic plane over $285^\circ < l < 65^\circ$, $|b| < 15^\circ$ on a bi-weekly cadence at 185--215\,MHz \citep[see Methods of][]{2023Natur.619..487H}; a full description will be released by Hurley-Walker et al., in prep. We also observed over June--September 2022 but made no detections in these data, as reported in \citetalias{2024MNRAS.535..909D}.}{The MWA, in its Phase~\textsc{ii} ``extended'' configuration \citep[][]{2018PASA...35...33W}, was used to scan the Galactic plane for transient radio sources under the `Galactic Plane Monitor', project code G0080. These observations were conducted over June--September\,2022, and June\,2024--March\,2025. Relevant to this work, the region $285^\circ < l < 65^\circ$, $|b| < 15^\circ$ was scanned on a bi-weekly cadence at 185--215\,MHz in both June--September\,2022 and 2024 \citep[see Methods of][a full description will be released by Hurley-Walker et al., in prep]{2023Natur.619..487H}. In this region, the root-mean-squared (RMS) noise level on the processed four-second timescale is $\sim$80\,mJy\,beam$^{-1}$. Amongst several detection methods (described in full by Horv\'ath et al., submitted, a convolution filter is employed to extract pulse-like signals. In June--September 2022, no pulses from \src{} were detected, but in June 2024, the source was blindly detected as a ${\sim}70$-s-wide pulse whose peak flux density was 665\,mJy.}

Subsequent to its redetection, we commenced a Director's Discretionary Time (DDT) campaign under project code D0042, to obtain further observations, adjusting the frequency from 200.32\,MHz to 184.96\,MHz for some of the observations to attempt to improve measurements of its DM.
Time of arrival (ToA) predictions for the DDT observations were made using a preliminary ephemeris derived from the GPM detections, which, by virtue at being at the same or similar frequencies, absorbed the delays due to dispersion and scattering into the `PEPOCH' term.

@@

We identified two faint pulses (4--5 $\sigma$; see Table~\ref{tbl:toas}) in the observation taken on 2024-10-04. However, observations taken on 2024-10-12 did not yield any pulses (at 3-$\sigma$ level). We inspected the data around the expected time of arrival (using Equation~\ref{eqn:timing}) but did not find any pulses. The instantaneous noise level, per 10\,s integration time bin, around the expected time of arrivals is close to 1\,mJy\,beam$^{-1}$. If pulse-to-pulse flux density variations, similar to ones observed in MWA data (Table~\ref{tbl:toas}), also occur at L/S band, then even a factor 1.5--2 variation can be sufficient to explain the detection of pulses on 2024-10-04 and the absence of pulses on 2024-10-12, given that the source is faint at higher frequencies.

@@

For the MeerKAT pulse with two components, we fit only the brighter component.
\revision{}{We checked the higher time resolution PTUSE data for finer substructure, but did not find anything smaller than the components that can be seen in \Fig~\ref{fig:pulsestack}.}

@@

These ToAs were then barycentred and fit for both period and DM, resulting in the values given in \Tab~\ref{tbl:ephemeris}.
The residuals are presented in \Fig~\ref{fig:pulse_details} and \Tab~\ref{tbl:toas}, along with other fitted properties of the pulses.
\revision{We also tried fitting a timing model that included a spin-down parameter, and found that it was consistent with zero spin-down to within $1\sigma$, and so is not included in the ephemeris.}{We also tried fitting a timing model that included a spin-down/up parameter (fixing the DM to $733\,{\rm pc}\,{\rm cm}^{-3}$), and found that it was consistent with zero.
However, owing to the recent discovery that the LPT CHIME/ILT\,J1634+44 is spinning up \citep[$\dot{P} = (-9.03 \pm 0.11) \times 10^{-12}\,{\rm s}\,{\rm s}^{-1}$;][]{2025arXiv250705139D,2025arXiv250705078B}, we report the fitted period derivative value of $(-10 \pm 92) \times 10^{-12}\,{\rm s}\,{\rm s}^{-1}$ instead of only providing an upper limit for the spin-down.}

@@

The variability of the pulse brightness, coupled with the fact that none of our detections were simultaneous at multiple telescopes, prohibit a direct spectral index measurement.
We note, however, that the fitted peak flux densities of the ASKAP and MeerKAT pulses are broadly consistent with the MWA flux densities at nearby epochs if an approximate spectral index of $\alpha = -2.0$ is assumed.
The bottom panel of \Fig~\ref{fig:pulse_details} shows the peak flux densities scaled to 1\,GHz under this assumption, revealing a relatively slow change of the pulses' brightness, not unlike GLEAM-X\,J1627$-$5235 \citep{2022Natur.601..526H}.
The in-band spectral index measurement of $\alpha=-3.1$ reported for the 2023 ASKAP pulse \citepalias{2024MNRAS.535..909D}, if consistent across the timespan of our observations, implies a turnover or spectral curvature at intermediate frequencies \citep[perhaps comparable with GPM\,1839$-$10;][]{2023Natur.619..487H}.

@@

We conjecture that if \src{} is a WD pulsar, its intermittency may be a result of the spin-orbital resonance being near, but not exactly, a small-integer ratio, creating a aliased beating pattern on a time scale of months.
The same explanation may be applied to other LPTs with activity cycles that last on the order of months.
An alternative explanation is that the radio emission is drived by particle wind outflows from \revision{the}{a} main sequence companion.
Long-term monitoring of LPTs will reveal whether the intermittency is periodic, and thus help discriminate between these two scenarios.

@@

This research is supported by an Australian Government Research Training Program (RTP) Scholarship.

@@

% ADACS
The database of MWA candidates, and the webapp designed to facilitate manually inspecting them, were developed by Astronomy Data and Computing Services (ADACS) as part of its software support scheme.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.48\linewidth]{J1755_optical.pdf}
    \includegraphics[width=0.48\linewidth]{J1755_nir.pdf}
    \caption{The $1\sigma$ positional uncertainty of \src{} derived from the MeerKAT observation taken on MJD 60602 is shown as the green ellipses overlaid on the optical (z, i, r) Panoramic Survey Telescope and Rapid Response System \citep[PanSTARRS;][]{2010SPIE.7733E..0EK,2016arXiv161205560C} image (left panel), and the infrared (K, H, J) United Kingdom Infrared Telescope Deep Sky Survey \citep[UKIDSS;][]{2007MNRAS.379.1599L} image (right panel).}
    \label{fig:localization}
\end{figure*}
